---
title: "Phenotype QC"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'Phenotype.html'))})
output: 
  html_document:
   code_folding: hide
   toc: true
   toc_float: true
   number_sections: TRUE
---

<br><br>

```{r setup, include=FALSE}
WD="~/OneDrive - The University of Queensland/mater_projs/auti_proj/AutismSpeaks_SPARK/"
knitr::opts_knit$set(root.dir = WD)
knitr::opts_chunk$set(echo = FALSE)
library(png)
library(grid)
library(DT)
```

# Probands  {.tabset .tabset-fade .tabset-pills}
- Note that the measurement of sum and mean is similar, thus only sum is presented.
- Adjusted by **Age, Sex and 20PCs** and then rank-based invserse normalise transformed (RINT). **FSIQ** was not included as a covariate due to too many missing values.

<br>

```{r phenotype_summary}
sums <- read.table("outputs/phenotype_summary.txt", header = T)
datatable(sums, rownames = FALSE, filter="top", caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: center; color:black; 
        font-size:200% ;',"Phenotype Summary"), options = list(pageLength = 5, scrollX=T) )
```


```{r adj_phe, eval=FALSE, include=FALSE}
###process phenotypes
###############################################################################
##extract phenotype IDs
setwd("/QRISdata/Q3457/SPARK/pheno/SPARK_Collection_Version6")

library(data.table)
fam <- fread("../../genotypes/SPARK.WES1.release.2021_03.genotype.fam")
eurs <- fread("/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/EUR.id", header = F)
names(eurs) <- c("FID", "IID")
##extract probs
role_index <- fread("roles_index.csv")
probs <- role_index[proband == "TRUE"][,.(family_sf_id, subject_sp_id)]
length(intersect(probs$subject_sp_id,fam$V2)) ##8406

names(probs) <- c("FID", "IID")

probs_eur <- probs[IID %in% eurs$IID]
fwrite(probs, "/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/probs.id", sep = "\t", col.names = F)
fwrite(probs_eur, "/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/probs_eur.id", sep = "\t", col.names = F)

##extract unaff_sibs
sib <- fread("background_history_sibling.csv")
unaff_sib <- sib[asd == "FALSE"][,.(family_sf_id, subject_sp_id)]
length(intersect(unaff_sib$subject_sp_id,fam$V2)) ##2268

names(unaff_sib) <- c("FID", "IID")
unaff_sib_eur <- unaff_sib[IID %in% eurs$IID]

fwrite(unaff_sib, "/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/unaffected_sibs.id", sep = "\t", col.names = F)
fwrite(unaff_sib_eur, "/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/unaffected_sibs_eur.id", sep = "\t", col.names = F)


###############################################################################
##RINT phenotypes
setwd("/QRISdata/Q3457/SPARK/pheno")
library(data.table)
fam <- fread("/QRISdata/Q3457/SPARK/genotypes/SPARK.WES1.release.2021_03.genotype.fam")
probs <- fread("/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/probs.id", header = F)
sibs <- fread("/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/unaffected_sibs.id", header = F)
phe1 <- fread("RBS_R.txt")
phe2 <- fread("SCQ.txt")
pcs <- fread("/QRISdata/Q3457/SPARK/hrc_imputed/QCd/grm/1000G_SPARK_proj_pruned0.3.proj.eigenvec")[,c(1:22)]
names(pcs) <- c("FID", "IID", paste0("PC", 1:20))

phe1_2 <- phe1[,c(1:13, 17)][subject_sp_id %in% fam$V2]
phe1_2[,FID := fam[match(phe1_2$subject_sp_id, fam$V2), V1]] ##6347
phe1_probs <- phe1_2[subject_sp_id %in% probs$V2] ##5634

phe2_2 <- phe2[,c(1:17, 21)][subject_sp_id %in% fam$V2]
phe2_2[,FID := fam[match(phe2_2$subject_sp_id, fam$V2), V1]] ##7553
phe2_probs <- phe2_2[subject_sp_id %in% probs$V2] ##6688

##Note 1/0 intersecting with sibs ID

names(phe1_probs)[1] <- "IID"
phe1_probs <- pcs[phe1_probs, on = c("FID", "IID")]

xnam <- paste("PC", 1:20, sep="")
phe_adj1 <- phe1_probs[,c("FID", "IID", "Age_Months", "sex")]
for(ncol in 25:34){
  y <- colnames(phe1_probs)[ncol]
  dat <- phe1_probs[,c(..y, "Age_Months", "sex", ..xnam)]
  names(dat)[1] <- "Y"
  
  fmla <- as.formula(paste("Y ~ Age_Months + ", paste(xnam, collapse= "+") ))
  phe_resids <- resid(lm(fmla, data = dat, na.action=na.exclude))
  Tmp <- data.frame(phe1_probs[,c("FID", "IID", "sex")], phe_resids)
  for(i in levels(factor(Tmp$sex))){
    Tmp[Tmp$sex == i, "phe_resids"]<- qnorm((rank(Tmp[Tmp$sex == i,]$phe_resids,na.last="keep") - 0.5)/sum(!is.na(Tmp[Tmp$sex == i,]$phe_resids))) 
  }
  phe_rint <- Tmp[match(phe_adj1$IID, Tmp$IID), "phe_resids"]
  phe_adj1[,y := phe_rint]
  phe_adj1[,(ncol-20)] <- phe_rint
  names(phe_adj1)[(ncol-20)] <- y
}

fwrite(phe_adj1, file = "/QRISdata/Q3490/SPARK/phenotypes/RBS_R_probands_adjPCs.txt", sep = "\t", na = NA)




names(phe2_probs)[1] <- "IID"
phe2_probs <- pcs[phe2_probs, on = c("FID", "IID")]

xnam <- paste("PC", 1:20, sep="")
phe_adj2 <- phe2_probs[,c("FID", "IID", "Age_Months", "sex")]
for(ncol in 25:38){
  y <- colnames(phe2_probs)[ncol]
  dat <- phe2_probs[,c(..y, "Age_Months", "sex", ..xnam)]
  names(dat)[1] <- "Y"
  
  fmla <- as.formula(paste("Y ~ Age_Months + ", paste(xnam, collapse= "+") ))
  phe_resids <- resid(lm(fmla, data = dat, na.action=na.exclude))
  Tmp <- data.frame(phe2_probs[,c("FID", "IID", "sex")], phe_resids)
  for(i in levels(factor(Tmp$sex))){
    Tmp[Tmp$sex == i, "phe_resids"]<- qnorm((rank(Tmp[Tmp$sex == i,]$phe_resids,na.last="keep") - 0.5)/sum(!is.na(Tmp[Tmp$sex == i,]$phe_resids))) 
  }
  phe_rint <- Tmp[match(phe_adj2$IID, Tmp$IID), "phe_resids"]
  phe_adj2[,y := phe_rint]
  phe_adj2[,(ncol-20)] <- phe_rint
  names(phe_adj2)[(ncol-20)] <- y
}

fwrite(phe_adj2, file = "/QRISdata/Q3490/SPARK/phenotypes/SCQ_probands_adjPCs.txt", sep = "\t", na = NA)
```

## Not Adjusted {.tabset .tabset-fade .tabset-pills}
```{r plot_For_notadj, eval=FALSE, include=FALSE}
library(data.table)
library(ggplot2)
setwd("/scratch/user/uqywan67/spark/data")

phe_unadj <- fread("/QRISdata/Q3457/SPARK/pheno/RBS_R.txt")
fam <- fread("SPARK_imputed_autosome_geno.05_maf.01_hwe1e-6_info.3.fam")
probs <- fread("probs.id", header = F)

phe_unadj1 <- phe_unadj[subject_sp_id %in% fam$V2 & subject_sp_id %in% probs$V2]

vecs <- grep("sum", names(phe_unadj1))
phe_unadj2 <- phe_unadj1[,c(1, ..vecs)]
phe_unadj3 <- melt(phe_unadj2, id = "subject_sp_id")

sums <- phe_unadj3[,.(sum(!is.na(value)), length(levels(factor(value)))), by = "variable"]
names(sums) <- c("variable", "Counts", "levels")

p <- ggplot(phe_unadj3,aes(x = value)) + 
  geom_histogram(alpha = 0.8) +
  #geom_text( aes(label= paste0(sCounts, " (", levels, ")"), x = 0, y = 250), hjust = -0.05 ) +
  facet_wrap(~variable,scales = "free", nrow = 1) +     
  
  theme_bw() +
  theme(
    text = element_text(size=14, face="bold"),
    axis.title.x = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 16),
    legend.title = element_blank()
  ) +
  ylab("Frequency") 

ggsave(file.path("figures/RBS_R_probands_Notadj.png"), p, height = 5, width = 16, dpi = 300)



phe_unadjII <- fread("/QRISdata/Q3457/SPARK/pheno/SCQ.txt")
phe_unadjII1 <- phe_unadjII[subject_sp_id %in% fam$V2 & subject_sp_id %in% probs$V2]

vecs <- grep("sum", names(phe_unadjII1))
phe_unadjII2 <- phe_unadjII1[,c(1, ..vecs)]
phe_unadjII3 <- melt(phe_unadjII2, id = "subject_sp_id")

sums2 <- phe_unadjII3[,.(sum(!is.na(value)), length(levels(factor(value)))), by = "variable"]
names(sums2) <- c("variable", "Counts", "levels")

p2 <- ggplot(phe_unadjII3,aes(x = value)) + 
  geom_histogram(alpha = 0.8) +
  #geom_text( aes(label= paste0(sCounts, " (", levels, ")"), x = 0, y = 250), hjust = -0.05 ) +
  facet_wrap(~variable,scales = "free", nrow = 2) +     
  
  theme_bw() +
  theme(a
    text = element_text(size=14, face="bold"),
    axis.title.x = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 16),
    legend.title = element_blank()
  ) +
  ylab("Frequency") 

ggsave(file.path("figures/SCQ_probands_Notadj.png"), p2, height = 10, width = 16, dpi = 300)


sums_all <- rbind(sums, sums2)
fwrite(sums_all, file = "phenotype_summary.txt", sep = "\t")

```


### RBS-R 
```{r plot_rbs-r, out.width="100%", out.height="100%", dpi=300}
grid.raster(readPNG("figures/RBS_R_probands_Notadj.png"))
```

<br>



***


### SCQ
```{r plot_scq, out.width="100%", out.height="100%", dpi=300}
grid.raster(readPNG("figures/SCQ_probands_Notadj.png"))
```

<br>


***

<br>

## Adjusted {.tabset .tabset-fade .tabset-pills}

### RBS-R
```{r plot_rbs-r_adj, out.width="100%", out.height="100%", dpi=300}
grid.raster(readPNG("figures/RBS_R_probands_adj.png"))
```

<br>

***

### SCQ
```{r plot_scq_adj, out.width="100%", out.height="100%", dpi=300}
grid.raster(readPNG("figures/SCQ_probands_adj.png"))
```

<br>


***


